Tool-driven Revolutions in Archaeological Science

### 2018-09-09 analysis/paper/jcaa-paper.Rmd
correct end of the sentence?

While all of these four factors could be considered relevant to the tooling of ecology and archaeology, we focus here on the third one because of its relevance to archaeology. Archaeology and ecology each have a high diversity of data collection methods, and types of data analyses they conduct, with only a small overlap. What they share in common is working with the field-collected data on a computer to prepare it for publication. R is the most common open source and free ware data analysis tool available and therefore a good proxy for the adoption of open source scientific programming languages tools in a discipline. To investigate changes in the use of R in archaeology and ecology, we obtained reference lists from a sample of scholarly articles in the Web of Science database and examined patterns in the citation of the R program over time.

@@

The second detail is that most archaeology journals do not show any increase in the percentage of article citing R over time . The ecology journals shown in Figure \@ref(fig:ecology-citing-r) show clear upward trends of increasing proportions of articles citing R over time, but we do not see any clear trends in the archaeology articles considered together. Indeed, in several of the archaeological journals in our sample, the first citations of R only occured in the last two years. The sample sizes for the archaeology journals are too small to confidently infer any trend of over time, with the exception of the _Journal of Archaeological Science_ with `r how_many_articles_in_jas` articles. The inset plot in Figure \@ref(fig:rchaeology-journals-with-subplot) shows a statistical test of change over time for articles in the _Journal of Archaeological Science_. With a moderate r-squared value and low p-value for the linear model, we can conclude that there is some evidence for an increase in citations of R over time in this journal. This may indicate an increasing use of R among archaeologists, especially those working at the intersection of the natural and physical sciences. This may hint at the start of a trend similar to what we see in the ecology journals of a widespread

>>>>

While all of these four factors could be considered relevant to the tooling of ecology and archaeology, we focus here on the third one because of its relevance to archaeology. Archaeology and ecology each have a high diversity of data collection methods, and types of data analyses they conduct, with only a small overlap. What they share in common is working with the field-collected data on a computer to prepare it for publication. R is the most common open source and free ware data analysis tool available and therefore a good proxy for the adoption of open source scientific programming languages in a discipline. To investigate changes in the use of R in archaeology and ecology, we obtained reference lists from a sample of scholarly articles in the Web of Science database and examined patterns in the citation of the R program over time.

@@

The second detail is that most archaeology journals do not show any increase in the percentage of article citing R over time . The ecology journals shown in Figure \@ref(fig:ecology-citing-r) show clear upward trends of increasing proportions of articles citing R over time, but we do not see any clear trends in the archaeology articles considered together. Indeed, in several of the archaeological journals in our sample, the first citations of R only occured in the last two years. The sample sizes for the archaeology journals are too small to confidently infer any trend of over time, with the exception of the _Journal of Archaeological Science_ with `r how_many_articles_in_jas` articles. The inset plot in Figure \@ref(fig:rchaeology-journals-with-subplot) shows a statistical test of change over time for articles in the _Journal of Archaeological Science_. With a moderate r-squared value and low p-value for the linear model, we can conclude that there is some evidence for an increase in citations of R over time in this journal. This may indicate an increasing use of R among archaeologists, especially those working at the intersection of the natural and physical sciences. This may hint at the start of a trend similar to what we see in the ecology journals of a widespread adoption of R.
